### 2023-10-06 docs/paper.qmd
working on results

---
title: "Reputational Risk and Questionable Research Practices in Archaeological Science"
author: Ben Marwick
date: now
format: html
editor: source
execute: 
  warning: false
---

## Introduction

```{r}
library(tidyverse)
```

## Authenticate with Elsevier

```{r}
#| eval: false

#  This only works on campus or with VPN
library(rscopus)
options("elsevier_api_key" = "053e6f8f2c1abe6fcdc943ee66759e5e")
have_api_key()
```

## Get JASR data

```{r}
#| eval: false

# get a bunch of DOIs for J. Arch. Sci Reports
res = scopus_search(query = "ISSN(2352-409X)", 
                    max_count = 6000, # Total Entries are 3906
                    count = 25)
df = gen_entries_to_df(res$entries)
head(df$df)

full_text_jasr <- vector("list", length = nrow(df$df))

library(httr)

# get the full text for each DOI
for(i in 1:nrow(df$df)){

  the_doi <- 
    paste0("https://api.elsevier.com/content/article/doi/",
           df$df$`prism:doi`[i], 
           "?APIKey=",
           "053e6f8f2c1abe6fcdc943ee66759e5e",
           "&httpAccept=text/plain")

  result <- GET(url = the_doi)

  full_text_jasr[[i]] <- content(result)

  print(i)
  Sys.sleep(0.1)

}

ft_df <- 
  tibble(
    doi = df$df$`prism:doi`,
    full_text_jasr = full_text_jas
  )

# save this so we don't have to scrape again
saveRDS(ft_df,
        "data/full_text_jasr.rds")
```

## Get JASR data

```{r}
#| eval: false

# get a bunch of DOIs for J. Arch. Sci
res = scopus_search(query = "ISSN(0305-4403)", 
                    max_count = 5000, # Total Entries are 6114
                    count = 25)
df = gen_entries_to_df(res$entries)
head(df$df)

full_text_jas <- vector("list", length = nrow(df$df))

library(httr)

# get the full text for each DOI
for(i in 1:nrow(df$df)){

  the_doi <- 
    paste0("https://api.elsevier.com/content/article/doi/",
           df$df$`prism:doi`[i], 
           "?APIKey=",
           "053e6f8f2c1abe6fcdc943ee66759e5e",
           "&httpAccept=text/plain")

  result <- GET(url = the_doi)

  full_text_jas[[i]] <- content(result)

  print(i)
  Sys.sleep(0.1)

}

ft_df <- 
  tibble(
    doi = df$df$`prism:doi`,
    full_text_jas = full_text_jas
  )

# save this so we don't have to scrape again
saveRDS(ft_df,
        "data/full_text_jas.rds")

```

## Clean the JASR data and extract p-values

```{r}
#| eval: false

ft_df_jasr <- readRDS("data/full_text_jasr.rds")

ft_df_clean <- 
  ft_df_jasr %>% 
  filter(map_lgl(full_text_jasr, ~ .x %>%
                   is.character %>% 
                   all)) %>% 
  mutate(full_text_jasr = map(full_text_jasr, 
                             ~str_squish(str_to_lower(.x)))) %>% 
  mutate(ps = map(full_text_jasr, 
                  ~unlist(str_extract_all(.x,
                                          "p *= .{6}|p *< .{6}|p *> .{6}"))))  %>% 
  unnest(ps)

# is the p-value reported exactly or with an inequi?
ft_df_clean_reporting <- 
ft_df_clean %>% 
  mutate(reporting = ifelse(str_detect(ps, "<|>"), "inequ", "exact"))

# take a look at how tests are reported by grabbing the text immediately before the
# p-values
jas_archaeology_reports_full_text_ps_test_reporting <- 
  ft_df %>% 
  mutate(test_reports = map(full_text_jasr, 
                            ~unlist(str_extract_all(str_squish(.x), 
                            ".{50}p = .{50}|.{50}p < .{50}|.{50}p > .{50}")))) %>% 
  unnest(test_reports)  %>% 
  select(-full_text_jasr)

# clean the p-values and treat inqualitites
ft_df_ps_clean <- 
  ft_df_clean_reporting %>% 
  mutate(p_value = case_when(
    str_detect(ps, "e")       ~  0.00001,
    str_detect(ps, "<0.0001") ~  0.00005,
    str_detect(ps, "< 0.0001") ~ 0.00005,
    str_detect(ps, "<0.001") ~   0.0005,
    str_detect(ps, "< 0.001") ~  0.0005,
    str_detect(ps, "< 0.05") ~   0.025,
    str_detect(ps, "<0.05") ~    0.025,
    str_detect(ps, "< 0.01") ~   0.005,
    str_detect(ps, "<0.01") ~    0.005,
    TRUE ~ parse_number(ps)
  ))  %>% 
  select(-full_text_jasr)

# save this so we don't have to clean again
write.csv(ft_df_ps_clean,
          "data/jasr_p_values_only.csv")
```

## Clean the JAS data and extract p-values

```{r}
#| eval: false

ft_df_jas <- readRDS("data/full_text_jas.rds")

ft_df_clean <- 
  ft_df_jas %>% 
  filter(map_lgl(full_text_jas, ~ .x %>%
                   is.character %>% 
                   all)) %>% 
  mutate(full_text_jas = map(full_text_jas, 
                             ~str_squish(str_to_lower(.x)))) %>% 
  mutate(ps = map(full_text_jas, 
                  ~unlist(str_extract_all(.x,
                                          "p *= .{6}|p *< .{6}|p *> .{6}"))))  %>% 
  unnest(ps)

# is the p-value reported exactly or with an inequi?
ft_df_clean_reporting <- 
ft_df_clean %>% 
  mutate(reporting = ifelse(str_detect(ps, "<|>"), "inequ", "exact"))

# take a look at how tests are reported by grabbing the text immediately before the
# p-values
jas_archaeology_full_text_ps_test_reporting <- 
  ft_df %>% 
  mutate(test_reports = map(full_text_jas, ~unlist(str_extract_all(str_squish(.x), ".{50}p = .{50}|.{50}p < .{50}|.{50}p > .{50}")))) %>% 
  unnest(test_reports)  %>% 
  select(-full_text_jas)

# clean the p-values and treat inqualitites
ft_df_ps_clean <- 
  ft_df_clean_reporting %>% 
  mutate(p_value = case_when(
    str_detect(ps, "e")       ~  0.00001,
    str_detect(ps, "<0.0001") ~  0.00005,
    str_detect(ps, "< 0.0001") ~ 0.00005,
    str_detect(ps, "<0.001") ~   0.0005,
    str_detect(ps, "< 0.001") ~  0.0005,
    str_detect(ps, "< 0.05") ~   0.025,
    str_detect(ps, "<0.05") ~    0.025,
    str_detect(ps, "< 0.01") ~   0.005,
    str_detect(ps, "<0.01") ~    0.005,
    TRUE ~ parse_number(ps)
  ))  %>% 
  select(-full_text_jas)

# save this so we don't have to clean again
write.csv(ft_df_ps_clean,
          "data/jas_p_values_only.csv")

```

## Explore the distributions of p-values

```{r}
jasr_p_values_only <- 
  read_csv(here::here("data/jasr_p_values_only.csv")) %>% 
  filter(between(p_value, 0, 1))

jas_p_values_only  <- 
  read_csv(here::here("data/jas_p_values_only.csv")) %>% 
  filter(between(p_value, 0, 1))

both_journals <- 
bind_rows(jasr_p_values_only,
          jas_p_values_only) %>% 
  mutate(journal = case_when(
    str_detect(doi, "jas\\.") ~ "JAS",
    str_detect(doi, "jasrep\\.") ~ "JASR"
  ))  %>% 
  drop_na(journal)

both_journals_summary <- 
both_journals %>% 
  group_by(journal, doi) %>% 
  summarise(n_p_values_per_article = n()) %>% 
  ungroup(doi) %>% 
  mutate(n_p_values_total = n()) %>% 
  mutate(journal_n = paste0(journal, "\n(n = ", n_p_values_total, ")"))

# drop unusual articles that have >100 p-values
articles_with_more_than_100_pvalues <- 
  both_journals_summary %>% 
  filter(n_p_values_per_article > 100) %>% 
  pull(doi)

both_journals_subset <- 
  both_journals %>% 
  filter(!doi %in% articles_with_more_than_100_pvalues)

n_p_values_total_jasr <- 
  both_journals_subset %>% 
  filter(journal == "JAS") %>% 
  nrow()

n_p_values_total_jas <- 
  both_journals_subset %>% 
  filter(journal == "JASR") %>% 
  nrow()

n_papers_jasr <- 
 both_journals_subset %>% 
  filter(journal == "JASR") %>% 
  distinct(doi) %>% 
  nrow()

n_papers_jas <- 
 both_journals_subset %>% 
  filter(journal == "JAS") %>% 
  distinct(doi) %>% 
  nrow()

both_journals_summary <- 
both_journals_subset %>% 
  group_by(journal, doi) %>% 
  summarise(n_p_values_per_article = n()) %>% 
  ungroup(doi) %>% 
  mutate(n_p_values_total = n()) %>% 
  mutate(journal_n = paste0(journal, "\n(n = ", n_p_values_total, ")"))

mean_number_of_p_values_per_paper <- 
  both_journals_summary %>% 
  summarise(mean_n_p = mean(n_p_values_per_article))

ggplot(both_journals_summary) +
  aes(journal_n, 
      n_p_values_per_article 
      ) +
  geom_boxplot() +
  scale_y_log10(labels = scales::label_comma(drop0trailing = TRUE)) +
  theme_minimal(base_size = 14)  +
  labs(x = "",
       y = "Number of p-values per article")

ggsave(here::here("figures/summary-n-papers-and-n-p-values.png"),
       h = 5,
       w = 4,
       bg = "white")

```

We have `r n_p_values_total_jasr` p-values from `r n_papers_jasr` JASR papers and `r n_p_values_total_jas` p-values from `r n_papers_jas` from JAS papers.

## p-values over time

```{r}

both_journals_by_year <- 
both_journals_subset %>% 
  mutate(year = str_remove(doi, "10.1016/j.jas[a-z]{0,3}.")) %>% 
  mutate(year = parse_number(str_sub(year, 1, 4))) %>% 
  filter(year > 1900) %>% 
  group_by(year, journal, doi) %>% 
  summarise(n_p_values_per_article = n()) %>% 
  ungroup(doi) %>% 
  summarise(n_p_values_per_year = n(),
            mean_number_of_p_values_per_paper = mean(n_p_values_per_article))

ggplot(both_journals_by_year) +
  aes(year, 
      mean_number_of_p_values_per_paper,
      colour = journal) +
  geom_line(linewidth = 2) +
  theme_minimal(base_size = 12) +
  theme(legend.position = c(0.7, 0.85)) +
  labs(y = "Mean number of p-values per paper",
       colour = "")

ggsave(here::here("figures/summary-p-values-over-time.png"),
       h = 4,
       w = 6,
       bg = "white")

```

## Inspect the distributions of p-values

```{r}
# zoom facet
# from here: https://www.data-imaginist.com/2019/the-ggforce-awakens-again/

library(ggforce)

jas_exact_zoom_plot_data <- 
  both_journals_subset %>% 
  filter(journal == "JAS",
         reporting == "exact")

ggplot() +
  geom_histogram(aes(p_value),
                 mutate(jas_exact_zoom_plot_data,
                        z = FALSE),
                 bins = 30)  +
  geom_histogram(aes(p_value),
                 mutate(jas_exact_zoom_plot_data,
                        z = TRUE),
                   bins = 400) +   
  facet_zoom(
    xlim = c(0, 0.1), 
    ylim = c(0, 600),
    zoom.data = z,
    horizontal = FALSE) +
  theme(zoom.y = element_blank(), 
        validate = FALSE) +
  scale_x_continuous(breaks = pretty) +
  theme_light(base_size = 18) +
  ylab("frequency in JAS") +
  xlab("p-value (exactly reported in the article text)")

ggsave(here::here("figures/jas_p_value_zoom_distribution.png"),
       w = 14,
       h = 6,
       dpi = 600)
```

```{r}

jasr_exact_zoom_plot_data <- 
  both_journals_subset %>% 
  filter(journal == "JASR",
         reporting == "exact")

ggplot() +
  geom_histogram(aes(p_value),
                 mutate(jasr_exact_zoom_plot_data,
                        z = FALSE),
                 bins = 30)  +
  geom_histogram(aes(p_value),
                 mutate(jasr_exact_zoom_plot_data,
                        z = TRUE),
                   bins = 400) +   
  facet_zoom(
    xlim = c(0, 0.1), 
    ylim = c(0, 600),
    zoom.data = z,
    horizontal = FALSE) +
  theme(zoom.y = element_blank(), 
        validate = FALSE) +
  scale_x_continuous(breaks = pretty) +
  theme_light(base_size = 18) +
  ylab("frequency in JASR") +
  xlab("p-value (exactly reported in the article text)")

ggsave(here::here("figures/jasr_p_value_zoom_distribution.png"),
       w = 14,
       h = 6,
       dpi = 600)
```