### 2023-10-11 README.Rmd
update readme and add some files

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "Reputational Risk and Questionable Research Practices in Archaeological Science."
```

# The Socio-Politics of Knowledge Production in Archaeology and Anthropology Workshop, October 2023

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/benmarwick/wg-sociopolitics-of-knowledge-production-in-archaeology-workshop/master?urlpath=rstudio)

This repository contains the R code and data for my presentation (available at [google slides](https://docs.google.com/presentation/d/1y6KGr1tUUwqQNrJE6gQPFboaMPaHxQmfLx0tpld8fKQ/edit#slide=id.ged4dcd42fc_0_167)) in this workshop:

## Reputational Risk and Questionable Research Practices in Archaeological Science.

Abstract: Uneven access to methods and data is an important dimension of global inequality in archaeological knowledge production. For example, physical instrumentation, such as a mass spectrometer, is often expensive to operate and highly immobile, requiring researchers to pay per use, and travel to locations to analyse samples. Computational methods and digital data, on the other hand, because of their nonphysical nature, have high potential for rapid and low-cost distribution far beyond their points of origin. Despite the low barriers to disseminating these materials, sharing of computational methods and digital data remains relatively rare in archaeology. This represents an unrealised opportunity to improve access and participation for underserved scholars, who may lack personal connections to learn new methods via direct instruction, and to acquire data through private arrangements. In this paper I explore if reputational risk might explain some of the reluctance for archaeologists to share their materials. If researchers are concerned that sharing code and data may enable others to identify either intentional or accidental questionable research practices (QRPs) that would harm their reputation, then we might predict low levels of sharing to be accompanied by high levels of QRPs. Alternatively, if non-reputational factors such as a lack of awareness or time limitations better explain low levels of sharing, we might predict low levels of QRPs. I investigate QRPs by measuring the prevalence of p-hacking (manipulation of a data analysis to produce a statistically significant result) in 5000 journal articles. The results show a strong signal of p-hacking, supporting reputational risk as a factor in the persistence of inequalities in archaeological knowledge production. To address this I propose de-emphasising reputation as a currency in research evaluation, for example by avoiding use of prestige metrics such as impact factors and h-indices. I further propose the research compendium as a template for communicating archaeological research that includes methods and data in human- and machine-readable formats to equalise the dissemination of archaeological science.

### How to cite

Please cite this compendium as:

> Authors, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

## Contents

The **analysis** directory contains:

  - [:file\_folder: docs](/docs): Quarto source document
    for manuscript. Includes code to reproduce the figures and tables
    generated by the analysis. It also has a rendered version,
    `paper.html`, suitable for reading (the code is replaced by figures
    and tables in this file)
  - [:file\_folder: data](/data): Data used in the analysis.
  - [:file\_folder: figures](/figures): Plots and other
    illustrations

## How to run in your browser or download and run locally

This research compendium has been developed using the statistical programming language R. To work with the compendium, you will need
installed on your computer the [R software](https://cloud.r-project.org/) itself and optionally [RStudio Desktop](https://rstudio.com/products/rstudio/download/).

You can download the compendium as a zip from from this URL:
[master.zip](/archive/master.zip). After unzipping:
- open the `.Rproj` file in RStudio
- run `renv::restore()` to ensure you have the packages this analysis depends on

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.